---
title: "Strings and Factors"
output: github_document
---

```{r, include = FALSE, message = FALSE, warning = FALSE}
library(tidyverse)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

Load key packages.

```{r}
library(rvest)
library(p8105.datasets)
```

## Mostly use string vectors

```{r}
string_vec = c("my", "name", "is", "jeff")

# look for whether a pattern is present in the string, gives TRUE or FALSE, it is case sensitive
str_detect(string_vec, "jeff")
str_detect(string_vec, "e")
str_detect(string_vec, "Jeff")

# replacing string or substring with another string
str_replace(string_vec, "jeff", "Jeff")
str_replace(string_vec, "e", "E")

# removing part of a string
str_replace(string_vec, "e", "")
str_remove(string_vec, "e")
```


(copied from p8105.com)

```{r}
string_vec = c(
  "i think we all rule for participating",
  "i think i have been caught",
  "i think this will be quite fun actually",
  "it will be fun, i think"
  )

str_detect(string_vec, "i think")
str_detect(string_vec, "^i think") # start of the string
str_detect(string_vec, "i think$") # end of the string

# removed i think from the end of the string
str_remove(string_vec, "i think$")
```


Pumpkin themed example!

```{r}
string_vec = c(
  "Time for a Pumpkin Spice Latte!",
  "went to the #pumpkinpatch last weekend",
  "Pumpkin Pie is obviously the best pie",
  "SMASHING PUMPKINS -- LIVE IN CONCERT!!"
  )

# capital p or lowercase p is detected (use brackets)
str_detect(string_vec, "[Pp]umpkin")
# detect all uppercase or all lowercase pumpkin
str_detect(string_vec, "[Pumpkinpumpkin]")
```


Let's get a bit more complicated


```{r}
string_vec = c(
  '7th inning stretch',
  '1st half soon to begin. Texas won the toss.',
  'she is 5 feet 4 inches tall',
  '3AM - cant sleep :('
  )

# detects any number 0-9 and a-z or A-Z, number followed by space isn't included
str_detect(string_vec, "[0-9][a-zA-Z]")
```


```{r}
string_vec = c(
  'Its 7:11 in the evening',
  'want to go to 7-11?',
  'my flight is AA711',
  'NetBios: scanning ip 203.167.114.66'
  )

# detect 7 followed by some character followed by 1
str_detect(string_vec, "7.1")
# detect 7 followed by a special character followed by 1 (\\ used to indicate a special character)
str_detect(string_vec, "7\\.1")

```


## Factors

```{r}
vec_sex = factor(c("male", "male", "female", "female"))
vec_sex

as.numeric(vec_sex)

# order based on what value comes first (male in this case)
vec_sex = fct_inorder(vec_sex)
vec_sex
```







## Revisit NSDUH

Import NSDUH data from the web

```{r}
url = "https://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"

# pulls html
drug_use_html = read_html(url)
```


This is an "easy" case

```{r}
nsduh_df = 
  drug_use_html |> 
  html_table() |> 
  first() |> # only keep first table
  slice(-1) # remove first row
```

Now do the tidying ...

```{r}
marj_df = 
  nsduh_df |> 
  select(-contains("P Value")) |> # remove "P value" from strings
  pivot_longer(
    -State,
    names_to = "age_year", 
    values_to = "percent"
  ) |> 
  separate(age_year, into = c("age", "year"), sep = "\\(") |>
  mutate(
    year = str_remove(year, "\\)"),
    percent = str_remove(percent, "[a-c]$"),
    percent = as.numeric(percent)
  ) |> 
  filter(
    !(State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West")) # only keep states, remove total and regions
  )
```

Let's make a quick plot

```{r}
marj_df |> 
  filter(age == "12-17") |> 
  mutate(State = fct_reorder(State, percent)) |> 
  ggplot(aes(x = State, y = percent, color = year)) +
  geom_point() + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) # common to google how to rotate the x-axis labels
```



## Restaurant Inspections!

```{r}
data("rest_inspec")
```

This is a fun dataset..


```{r}
# number of each grade by borough
rest_inspec |> 
  group_by(boro, grade) |> 
  summarize(n = n()) |> 
  pivot_wider(
    names_from = grade, 
    values_from = n
  )
```


update the dataset

```{r}
# focus on 5 boroughs and grades A, B, and C
rest_inspec =
  rest_inspec |> 
  filter(
    grade %in% c("A", "B", "C"),
    boro != 0)
```

Look for pizza places

```{r}
# look at pizza places, this isn't correct because some pizza places are in all caps
rest_inspec |> 
  filter(str_detect(dba, "Pizza")) |> 
  group_by(boro, grade) |> 
  summarize(n = n()) |> 
  pivot_wider(
    names_from = grade, 
    values_from = n
  )
```


```{r}
# look at pizza places (lowercase and uppercase), gives all possible combinations for pizza
rest_inspec |> 
  filter(str_detect(dba, "[Pp][Ii][Zz][Zz][Aa]")) |> 
  group_by(boro, grade) |> 
  summarize(n = n()) |> 
  pivot_wider(
    names_from = grade, 
    values_from = n
  )
```

Let's make this table a plot instead

```{r}
# bar plot of grade breakdown for pizza places across boroughs
# fill better than color for bar plots
# alphabetical order is the default
rest_inspec |> 
  filter(str_detect(dba, "[Pp][Ii][Zz][Zz][Aa]")) |> 
  ggplot(aes(x = boro, fill = grade)) + 
  geom_bar()
```



```{r}
rest_inspec |> 
  filter(str_detect(dba, "[Pp][Ii][Zz][Zz][Aa]")) |> 
  mutate(boro = fct_infreq(boro)) |> # order of frequency
  ggplot(aes(x = boro, fill = grade)) + 
  geom_bar()
```


This doesn't work as intended...

```{r}
rest_inspec |> 
  filter(str_detect(dba, "[Pp][Ii][Zz][Zz][Aa]")) |> 
  mutate(
    boro = fct_infreq(boro),
    boro = str_replace(boro, "Manhattan", "The City")) |> # back in alphabetical order
  ggplot(aes(x = boro, fill = grade)) + 
  geom_bar()
```


... but this does:

```{r}
rest_inspec |> 
  filter(str_detect(dba, "[Pp][Ii][Zz][Zz][Aa]")) |> 
  mutate(
    boro = fct_infreq(boro),
    boro = fct_recode(boro, "The City" = "Manhattan")) |> # this works for renaming a label
  ggplot(aes(x = boro, fill = grade)) + 
  geom_bar()
```



